---
title: 'Statistical Inference Course Project Part 2'
author: "Muhammad Imran Saeed"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
```
##  For this part of the project, we are going to analyze the ToothGrowth data in the R datasets package.

### Tasks to be perfromed for the assignment:
1. Load the ToothGrowth data and perform some basic exploratory data analyses.
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering).
4. State your conclusions and the assumptions needed for your conclusions.

###  Installation and setup.
```{r package setup and installation,include=FALSE}
listOfPackages <- c("ggplot2","plyr","tinytex")
newPackages <- listOfPackages[!(listOfPackages %in% installed.packages()[,"Package"])]

if(length(newPackages)) {
	message(sprintf("Going to install package(s)[ %s ]\n", newPackages))
	install.packages(newPackages) 
} else {
	message("All packages are already installed. Will skip install pahse")
}


## Load all of the package needed for this project.
for(package in listOfPackages){
  library(package, character.only = TRUE)
}
knitr::opts_chunk$set(size='\\small')
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(fig.align='center')
```

##### Loading and preprocessing the data
```{r load data}
data("ToothGrowth")

```
The data shows the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice (OJ) or ascorbic acid(VC) (a form of vitamin C).

The variable names and the structure of the file are given by
```{r describe data, paged.print=TRUE}
# Query names from data set
names(ToothGrowth)
# Query structure of data set
str(ToothGrowth)
# Check sample data limited to 4 rows to ensure maximum of 6 pages
head(ToothGrowth,4)
# Create needed columns for analysis.
ToothGrowth$dosage <- as.factor(ToothGrowth$dose)
```
Summary of the data.
```{r}
summary(ToothGrowth)
```
### Tooth Length to Delivery Method
```{r tooth-length-relative-to-delivery-method, echo = TRUE}
tg <- ToothGrowth
levels(tg$supp) <- c("Orange Juice", "Ascorbic Acid")
gLenSupp <- ggplot(data = tg, aes(x = supp, y = len)) + 
    geom_boxplot(aes(fill = supp)) +
    xlab("Delivery Method") +
    ylab("Tooth Length") +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    ggtitle("Tooth Length as a Function of Delivery Method")
print(gLenSupp)
```
**Observation**
Above indicates using OJ as delivery method,independent of dosage, has favorable effect on tooth growth than ascorbic acid.
### Tooth Length to Dosage Level
```{r tooth-length-relative-to-dosage-level, echo = TRUE}
gLenDose <- ggplot(data = ToothGrowth, aes(x = factor(dose), y = len)) + 
    geom_boxplot(aes(fill = factor(dose))) +
    xlab("Dosage Level (mg/day)") +
    ylab("Tooth Length") +
    guides(fill=guide_legend(title="dose")) +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    ggtitle("Tooth Length as a Function of Dosage Level")
print(gLenDose)
```

**Observation**

Above indicates higher dosage of VC irrespective of deliver method have favorable outcome on tooth growth. The best results are achieved by high dosage i.e 2mg/day followed by dosage in descending order(1mg ,0,5 mg/ day).

### Tooth Length to Delivery Method and Dosage Level
```{r tooth-length-relative-to-delivery-method-and-dosage-level, echo = TRUE}
tg <- ToothGrowth
levels(tg$supp) <- c("Orange Juice", "Ascorbic Acid")
gLenSuppDose <- ggplot(data = tg, aes(x = supp, y = len)) + 
    geom_boxplot(aes(fill = supp)) +
    facet_wrap(~ dose) +
    xlab("Delivery Method") +
    ylab("Tooth Length") +
    guides(fill=guide_legend(title="supp")) +
    theme(plot.title = element_text(size = 14, hjust = 0.5, vjust = 0.5),
          axis.text.x = element_text(angle = 45,
                                     hjust = 0.5,
                                     vjust = 0.5,
                                     margin = margin(b = 10))) +
    ggtitle("Tooth Length as a Function of Delivery Method and Dosage Level")
print(gLenSuppDose)
```

**Observation**
Above chart indicates higher dosage i.e mg/day is more effective than the lower dosage(s); however, all delivery methods have similar effectiveness.
```{r confidence-intervals-for-dosage, echo = TRUE}

Lower <- subset(tg, dose %in% c(0.5, 1.0))
Middle <- subset(tg, dose %in% c(0.5, 2.0))
Upper <- subset(tg, dose %in% c(1.0, 2.0))

```

Perform T test on basis of dosage for Lower.
```{r t-test-lower-dosage, echo = TRUE}

t.test(len ~ dose, paired = F, var.equal = F, data = Lower)

```

Perform T test on basis of dosage for Middle.
```{r t-test-middle-dosage, echo = TRUE}

t.test(len ~ dose, paired = F, var.equal = F, data = Middle)

```

Perform T test on basis of dosage for Upper.
```{r t-test-upper-dosage, echo = TRUE}
t.test(len ~ dose, paired = F, var.equal = F, data = Upper)
```


Perform T test on basis of supply
```{r t-test-for-supply, echo = TRUE}
t.test(len ~ supp, paired = F, var.equal = F, data = tg)
```

**Conclusions and the assumptions needed for your conclusions**

1. Supplement type has no effect on tooth growth as we can’t reject null hypothesis here
2. Increasing dosage results in positive tooth growth, the clear growth of confidence interval allow us to reject null hypothesis

